\section{Proving Techniques}

The properties that have been proven can be divided into three types: Preconditions, Invariants and Post-conditions \cite{meyer_1992}. Preconditions are properties that must be true before a function is called, post-conditions must be true after a function is called, and invariants are properties that must be true for all values of a certain type. In this section it will be shown that these three types of properties each have their own way to be proven in Agda.

\subsection{Finding properties to prove}\label{finding_props_to_prove}
\cite{BREITNER2021} presents multiple techniques to find properties to prove from Haskell code. A few of those techniques have been selected that were deemed useful for this paper:
\begin{enumerate}
	\item Define an invariant property when there are types whose correctness depend on invariants
		\cite[p. 7]{BREITNER2021}
	\item Define a post-condition property by deriving a definition directly from the test suite
		\cite[p. 9]{BREITNER2021}
	\item Define a precondition when there is a risk of numeric overflow, or switch to using unbounded integers instead. 
		\cite[p. 9]{BREITNER2021}
	\item Define a post-condition when there are type classes which come with laws that all instances of the type class should satisfy
		\cite[p. 10]{BREITNER2021}
\end{enumerate}
Finally, one technique that was not mentioned in \cite{BREITNER2021} was used:
\begin{enumerate}
	\setcounter{enumi}{4}
	\item Define a precondition if it is required to make the function total, as was described in section \ref{total-functions}
\end{enumerate}


\subsection{Properties to prove}\label{props_to_prove}

Instead of defining a precondition when there is a risk of numeric overflow as described in technique 3, we switch to using unbounded integers in this papers' implementation of the QuadTree library. This is the same decision that \cite{BREITNER2021} made. Furthermore, the test suite of the QuadTree library consists only of tests that test the type class laws, so no additional properties could be derived from technique 2. Using technique 1, 4, and 5, the following properties of the QuadTree library were derived: 

\setlength{\parindent}{0pt} \setlength{\parskip}{5pt}
\textbf{Invariants of a QuadTree:}
\begin{itemize}
    \item Depth invariant: The depth of a QuadTree must be less than or equal to \linebreak $\left \lceil{log_{2}( max(width, height))}\right \rceil$. This is to ensure that there is exactly one value at each location. (Technique 1)
    \item Compression invariant: No node can have four leaves that are identical. These need to be fused into a single leaf quadrant. This is needed to keep the QuadTree fast and space efficient. (Technique 1)
\end{itemize} 
\textbf{Preconditions of a QuadTree:}
\begin{itemize}
    \item When calling \verb|atLocation|, \verb|getLocation|, \verb|setLocation| or \verb|mapLocation|, the location must be inside of the QuadTree. (Technique 5)
    \item When calling \verb|lensLeaf|, the quadrant needs to have a depth of zero (i.e.  it must be a leaf) (Technique 5)
    \item When calling \verb|lensA/B/C/D|, the quadrant needs to have a depth that is greater than zero (i.e.  it must not be a leaf) (Technique 5)
\end{itemize} 
\textbf{Post-conditions of a QuadTree:}
\begin{itemize}
    \item The lenses returned by all the lens functions satisfy the lens laws: \cite{lens} (Technique 4)
        \begin{itemize}
            \item \verb|view l (set l v s) = v| (Setting and then getting returns the value)
            \item \verb|set l (view l s) s = s| (Setting the value to what it already was doesn't change anything)
            \item \verb|set l v2 (set l v1 s) = set l v2 s| (Setting a value twice is the same as setting it once to the second value)
        \end{itemize}
    \item The functor implementations for Quadrant and QuadTree satisfy the functor laws (Technique 4)
        \begin{itemize}
            \item \verb|fmap id = id| (Identity law)
            \item \verb|fmap (f . g) == fmap f . fmap g| (Composition law)
        \end{itemize}
    \item The foldable implementation returns an output of the correct length (Technique 4)
        \begin{itemize}
            \item \verb|length quadtreeFoldable vqt = width * height|
        \end{itemize}
    \item The foldable implementation satisfies the foldable-functor law (Technique 4)
        \begin{itemize}
            \item \verb|foldMap f = fold . fmap f|
        \end{itemize}
\end{itemize}
	
\setlength{\parindent}{15pt} \setlength{\parskip}{0pt}
\subsection{Techniques to prove invariants} \label{prove_invariants}
Invariants are proven by creating a new datatype with one constructor, which takes the original datatype and a proof for all the invariants. As a simple example, this datatype represents a natural number with the invariant that it is greater than 5.
\begin{minted}{agda}
data GreaterThanFive : Set where
  CGreaterThanFive : (n : Nat) -> { .( IsTrue (n > 5) ) } -> GreaterThanFive
\end{minted}
The proof is marked as implicit \{\} so that it is removed when compiled to Haskell, and it is marked as as irrelevant .() so that will not interfere when proving post-conditions later. An irrelevant value means that the actual value of the proof does not matter, only its existence does.

Using this technique, the datatype for a compressed quadrant with a certain maximum depth is: 
\begin{minted}{agda}
data VQuadrant (t : Set) {depth : Nat} : Set where
  CVQuadrant : (qd : Quadrant t) 
            -> {.(IsTrue (depth qd <= depth && isCompressed qd))} 
            -> VQuadrant t {depth}
\end{minted}
The datatype for a valid QuadTree is defined very similarly. Agda2hs flawlessly compiles this to the following Haskell code, where the proof is erased:
\begin{minted}{haskell}
data VQuadrant t = CVQuadrant (Quadrant t)
\end{minted}

The advantage of making a new wrapper datatype over adding the proofs to the original datatype is that if the original datatype has multiple constructors, functions that use the proof do not need to be split into multiple cases (one for each constructor). The disadvantage is that this additional wrapper type is visible when compiled to Haskell. To avoid this, it is possible to create an additional function for all public functions. This function then takes the invariance proof as a precondition, and calls the original function with the wrapper type. 

\subsection{Techniques to prove preconditions}
In this section, two techniques to prove preconditions are presented.

\subsubsection{Using an implicit argument} \label{using_implicit}
When using the implicit argument technique, preconditions are proven by adding the proofs as implicit arguments to the function.
As a simple example, this function takes a natural number that must be greater than 5.
\begin{minted}{agda}
takesGtFive : (n : Nat) -> { .( IsTrue (n > 5) ) } -> ?
\end{minted}
As with invariants, the proof is marked as implicit and irrelevant. 

Using this technique, a precondition can be used to ensure that the location given to the \verb|getLocation| function must be inside the QuadTree:
\begin{minted}{agda}
-- Function that checks if a location is inside a given QuadTree
isInsideQuadTree : (Nat × Nat) -> QuadTree t -> Bool
isInsideQuadTree (x , y) (Wrapper (w , h) _) = x < w && y < h

getLocation : (loc : Nat × Nat) -> (qt : QuadTree t) 
    -> {.( IsTrue (isInsideQuadTree loc qt) )} -> t
\end{minted}
After being compiled with Agda2hs, the precondition is removed from the function, just like with invariants.
\begin{minted}{haskell}
getLocation :: (Nat, Nat) -> QuadTree t -> t
\end{minted}

\subsubsection{Using a datatype with invariants}
Another technique to prove preconditions is by passing in a datatype with an invariant, as was used in section \ref{prove_invariants}. The simple example from \ref{using_implicit} would then be written like this, using the type defined in section \ref{prove_invariants}:
\begin{minted}{agda}
takesGtFive : (n : GreaterThanFive) -> ?
\end{minted}

For the QuadTree verification, this technique was used to encode the maximum depth properties of the lens functions, using the same datatype that was defined for the invariants.
\begin{minted}{agda}
lensLeaf : Lens (VQuadrant t {0}) t
lensA : {dep : Nat} 
    -> Lens (VQuadrant t {S dep}) (VQuadrant t {dep})
\end{minted}

\subsubsection{Comparison}
The advantages of using implicit arguments is that it is not necessary to define a separate datatype, and that the precondition can be dependent on more than one parameter of the function. On the other hand, the advantages of using a datatype with an invariant is that the defined function are cleaner and more compact. It is then also possible to use the type as a parameter for another type, like it is used in \verb|lensLeaf| and \verb|lensA|. It also allows for cleaner reuse of the property, as it does not need to be repeated each time it is used.

\subsection{Techniques to prove post-conditions}
Post-conditions are proven as separate functions. As a simple example, this is a proof that this function returns a number greater than 5.
\begin{minted}{agda}
gt5 : Bool -> Nat
gt5 _ = 42

gt5-is-gt5 : (b : Bool) -> IsTrue (gt5 b > 5)
gt5-is-gt5 b = IsTrue.itsTrue
\end{minted}

For the QuadTree verification, this technique was used to verify the lens laws of all the lenses defined in the implementation. For example, this is the proof that the ViewSet law holds for \verb|lensLeaf|.
\begin{minted}{agda}
ValidLens-Leaf-ViewSet : 
    -> (v : t) (s : VQuadrant t {0}) 
    -> view (lensLeaf {t}) (set (lensLeaf {t}) v s) ≡ v
ValidLens-Leaf-ViewSet v (CVQuadrant (Leaf x)) = refl
\end{minted}

When proving preconditions and invariants, these properties have to be marked as irrelevant. This is to ensure that when proving that two function calls are equal, one does not need to show that the proofs of the preconditions and invariants are equal, since the actual value of the proofs is irrelevant.

\subsection{Results} \label{results}
All of the properties mentioned in section \ref{props_to_prove} have been successfully proven. Most of these proofs have not been shown in the paper, since they are available in the source code. The amount of lines of code that this took is shown in figure \ref{division}.  This counts all non-empty lines. The verification took about 3 times more lines of code than the implementation. While this comparison is an indication, this should not be taken to mean that the verification took 3 times as much effort, as the information density of the implementation and proofs is different.

In reality, the implementation took approximately one full-time week, while the verification took approximately five full-time weeks. This too should not be taken to mean the verification took 5 times as much effort, as this number may be biased by the fact that the implementation was just a translation from Haskell.

During the verification phase, one bug was found in this papers implementation of QuadTree that was accidentally introduced during the translation to Agda. This was not caught by the tests, though this may be because the tests on the foldable implementation are very limited.

Whether the verification is worth the time spent, depends on the situation. For example, in a situation where even one small error could bring down an airplane, this is clearly worth it. However, in many common situations, this verification may not be worth it.

\begin{figure}[H]
	\begin{tikzpicture}[scale=0.8]
		\pie[sum=auto]
		{757/Implementation,
			545/General Proofs,
			789/Foldable Proofs,
			809/Lens Proofs,
			215/Functor Proofs}
	\end{tikzpicture}
	\caption{Division of lines of code}
	\label{division}
\end{figure}